---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
library(dplyr)
data=read.csv("activity.csv",colClasses=c("numeric","Date","numeric"))
data_tb=tbl_df(data)


```

## What is mean total number of steps taken per day?

```{r}
pta1=filter(data_tb,!is.na(steps))
pta2=group_by(pta1,date)
pta3=summarise(pta2,steps=sum(steps))
pta4=filter(pta3,!is.na(steps))
barplot(pta4$steps,names.arg=pta4$date)

```

The mean total number of steps taken is `r mean(pta4$steps)` and the median total number of steps taken is `r median(pta4$steps)`.

## What is the average daily activity pattern?

```{r}
ptb1=group_by(data_tb,interval)
ptb2=filter(ptb1,!is.na(steps))
ptb3=summarise(ptb2,ave=mean(steps))
ptb3$interval=as.numeric(as.character(ptb3$interval))
ptb4=ptb3[order(ptb3$interval),]
plot(ptb4$interval,ptb4$ave,type='l', xlab='Interval',ylab='Average number of steps')
```

The interval with the most number of steps is interval `r as.numeric( ptb4[ptb4$ave==max(ptb4$ave),1])`.

## Imputing missing values
The number of rows with missing values is: `r nrow(filter(data_tb,is.na(steps)))`.  
Strategy: Use mean for the 5 min interval

```{r}
ptc1=data_tb
for (i in 1:nrow(ptc1)){
    if (is.na(ptc1[i,1])){
        a=filter(ptb4,interval==as.numeric(ptc1[i,3]))
        ptc1[i,1]=a[1,2]
    }
}
ptc2=group_by(ptc1,date)
ptc3=summarise(ptc2,steps=sum(steps))
barplot(ptc3$steps,names.arg=ptc3$date)
```

The mean total number of steps taken is `r mean(ptc3$steps)` and the median total number of steps taken is `r median(ptc3$steps)`.  
There is no change in the mean number of steps but there is a slight increase in the median.  
There is no significant impact as the missing values are estimated from the data itself.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
ptd1=ptc1
ptd2=mutate(ptd1,day=weekdays(date))
ptd3a=mutate(filter(ptd2,day=="Saturday"|day=="Sunday"),wkday="Weekend")
ptd3b=mutate(filter(ptd2,day!="Saturday"|day=="Sunday"),wkday="Weekday")
ptd4a=group_by(ptd3a,interval)
ptd4b=group_by(ptd3b,interval)
ptd5a=summarise(ptd4a,ave=mean(steps))
ptd5b=summarise(ptd4b,ave=mean(steps))
plot(ptd5b$interval,ptd5b$ave,type='l', xlab='Interval',ylab='Average number of steps', main='Weekday')
plot(ptd5a$interval,ptd5a$ave,type='l', xlab='Interval',ylab='Average number of steps', main='Weekend')
```
